manuscript_figures

### 2024-07-06 scripts/manuscript_figures.Rmd
updasted HPI and production plots

# Calculate fish biomass 'W' using the length-weight relationship
PB_filtered$biomass <- with(PB, coef.a * (actual_size)^coef.b)

# Divide biomass by 80 m to get g/m2
#calculate biomass/m2
PB$biomass.g.m2 <- PB$biomass / 80

# Constants from the P/B ratio equation
coefficient <- 2.64
exponent <- -0.35

# Calculate P/B ratio, the predictive equation is P/B (per year) = 2.64*Wmat^-035 , where Wmat is weight-at-maturity (Randall and Minns 2000)
PB$P_B_ratio <- coefficient * PB$biomass.g.m2^exponent

# Calculate Habitat Productivity Index (HPI), habitat productivity index (HPI), a proposed measure of habitat suitability and productive capacity, can be calculated as the product of average fish biomass and the species P/B coefficient (HPI = B × P/B) (Randall and Minns 2000)
PB$HPI <- PB$biomass.g.m2 * PB$P_B_ratio

# Aggregate by Transect, Site, Year and calculate mean HPI
hpi_summary <- PB %>%
  group_by(TRANSECT, SITE, YEAR) %>%
  summarise(mean_HPI = mean(HPI, na.rm = TRUE))

# Make faceted line plot of HPI index over the years
hpi.plot < - ggplot(hpi_summary, aes(x = interaction(TRANSECT, SITE, sep = " - "), y = mean_HPI, fill = as.factor(YEAR))) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  labs(x = "Transect - Site", y = "Mean Habitat Productivity Index (HPI)", fill = "Year") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1));hpi.plot

>>>>

# Calculate fish biomass 'W' using the length-weight relationship
PB_filtered$Wmat <- PB_filtered$coef.a * (PB_filtered$Lmat^PB_filtered$coef.b)

# Constants for the PB.ratio calculation
coefficient <- 2.64
exponent <- -0.35

# Calculate PB.ratio column
PB_filtered$PB.ratio <- coefficient * (PB_filtered$Wmat ^ exponent)

# Calculate fish biomass
PB_filtered$biomass <- PB_filtered$Wmat * PB_filtered$COUNT

# Calculate biomass per m²
PB_filtered$biomass_per_m2 <- PB_filtered$biomass / 80

# Plotting using ggplot2
production.plot <- ggplot(PB_filtered, aes(x = YEAR, y = PB.ratio, fill = SP_CODE)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Species",
       y = "P/B",
       fill = "Species") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),  # Rotate x-axis labels if needed
        axis.title.x = element_text(vjust = -0.5, size = 18),  # Adjust x-axis title position
        axis.title.y = element_text(vjust = 1.5, size = 18),
        axis.text.y = element_text(size = 18),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 18)) + # Adjust y-axis title position
  scale_y_continuous(limits = c(0, 4), expand = c(0, 0));production.plot  # Set y-axis limits and remove expansion

# Calculate Habitat Productivity Index (HPI), habitat productivity index (HPI), a proposed measure of habitat suitability and productive capacity, can be calculated as the product of average fish biomass and the species P/B coefficient (HPI = B × P/B) (Randall and Minns 2000)
# Calculate HPI
HPI_calculated <- PB_filtered %>%
  group_by(YEAR, SITE, TRANSECT) %>%
  summarise(HPI = sum(biomass_per_m2 * PB.ratio, na.rm = TRUE),
            .groups = 'drop')

# We add a small constant (1e-6) to avoid issues with log(0)
HPI_calculated <- HPI_calculated %>%
  mutate(log_HPI = log(HPI + 1.1))

# Now plot with the log-transformed HPI
species.hpi.plot <- ggplot(HPI_calculated, aes(x = SITE, y = log_HPI, fill = factor(YEAR))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Site", 
       y = "Log Habitat Productivity Index", 
       fill = "Year") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18),
        axis.text.y = element_text(size = 18),
        axis.title.y = element_text(size = 18),
         axis.title.x = element_text(size = 18),
        legend.text = element_text(size = 16),
        legend.title = element_text(size = 18)) +
  scale_y_continuous(expand = expansion(mult = c(0, .05)),
    labels = scales::label_number(accuracy = 0.1)) +
  coord_cartesian(clip = "off"); species.hpi.plot

# Plot two plots together using patchwork for production and HPI
```{r}
# Combine the plots with patchwork for all years of kelp and rate of change
production.hpi <- production.plot | species_hpi.plot +
  plot_annotation(tag_levels = 'A') + 
  theme(plot.tag = element_text(size = 20, face = "bold"));production.hpi
